---
title: "Additional OTU table analyses"
author: "Alex Linz"
date: "Monday, February 08, 2016"
output: html_document
---


```{r, CoV:epilimnia, echo=FALSE, warning=FALSE, message=FALSE}
#Function to calculate coefficient of variance (standard deviation/mean)
CoV <- function(vector){
  sample <- as.numeric(vector)
  m <- mean(sample)
  s <- sd(sample)
  return(s/m)
}

#Sort epilimnion clades by abundance
sorted_epi_clade_table <- epi_clade_table2[match(names(epi_clade_sums), rownames(epi_clade_table2)),]
#Calculate CoV
Cov_epi <- apply(sorted_epi_clade_table, 1, CoV)

#Make dataframe of top 20 clades by abundance for plotting
plot_cov_epi <- data.frame(names(Cov_epi)[1:20], Cov_epi[1:20])
colnames(plot_cov_epi) <- c("Clade", "CoV")
plot_cov_epi$Clade <- factor(plot_cov_epi$Clade, levels=plot_cov_epi$Clade)
ggplot(data=plot_cov_epi, aes(x=Clade, y=CoV)) + geom_bar(stat = "identity", fill = "grey", colour="black") + labs(x = NULL, y = "CoV in Epilimnia") + theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 12, colour = "black"), axis.title = element_text(size = 12, vjust=2), axis.text.y = element_text(colour="black")) + scale_y_continuous(expand = c(0,0)) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black"))
```

####Figure S5. Coefficient of Variance for the top 20 clades in epilimnia. A low CoV indicates low varability in abundance, while a high CoV indicates high variability in abundance. In epilimnia, some of the most abundant clades (Pnec, acI-B, betI-A, and bacI-A) have low CoV values, while *Cyanobacteria*, gamI, and *Geothrix* have high CoV values, indicating more variability in their abundance measurements. 
***
```{r, CoV: hypolimnia, echo=FALSE, warning=FALSE, message=FALSE}
#Repeat above analysis for hypo
sorted_hypo_clade_table <- hypo_clade_table2[match(names(hypo_clade_sums), rownames(hypo_clade_table2)),]
Cov_hypo <- apply(sorted_hypo_clade_table, 1, CoV)
plot_cov_hypo <- data.frame(names(Cov_hypo)[1:20], Cov_hypo[1:20])
colnames(plot_cov_hypo) <- c("Clade", "CoV")
plot_cov_hypo$Clade <- factor(plot_cov_hypo$Clade, levels=plot_cov_hypo$Clade)
ggplot(data=plot_cov_hypo, aes(x=Clade, y=CoV)) + geom_bar(stat = "identity", fill = "grey", colour="black") + labs(x = NULL, y = "CoV in hypolimnia") + theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 12, colour = "black"), axis.title = element_text(size = 12, vjust=2), axis.text.y = element_text(colour="black")) + scale_y_continuous(expand = c(0,0)) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black"))
```

####Figure S6. Coefficient of variance for the top 20 clades in hypolimnia (mixing dates and polymictic lake samples removed). A low CoV indicates low varability in abundance, while a high CoV indicates high variability in abundance. In epilimnia, some of the most abundant clades (Pnec, unclassified Bacteria, Geothrix, and unclassified Bacteroidetes) have low CoV values, while gamIII-A has a very high CoV value, indicating more variability in its abundance measurements.   
***
```{r, CoV:hypolimnion+mixing, echo=FALSE, warning=FALSE, message=FALSE}
#Compute difference in CoV when mixing is included vs when it is not included in the hypolimnion analysis.

#Make clade table with mixing dates removed
hypo_clade_table <- bog_subset("TBH|SSH|NSH|MAH|HKH", clade_table)
hits <- match(mixes, substr(colnames(hypo_clade_table), start = 1, stop = 10))
hypo_clade_table <- hypo_clade_table[,-hits]


#Remove taxonomic level tags at the beginning of some names (present in Greengenes classified clades)
rownames(hypo_clade_table) <- gsub("[kpcofg]__|\\[|\\]|Bacteria;", "", rownames(hypo_clade_table))

#Keep only clades that are completely classified
splitnames<- strsplit(rownames(hypo_clade_table), ";")
clades_only <- c()
for(i in 1:length(splitnames)){
  clades_only[i] <- splitnames[[i]][5]
}

hypo_clade_table2 <- hypo_clade_table[which(clades_only != "unclassified" & is.na(clades_only) == F),]
hypo_clade_table2 <- reduce_names(hypo_clade_table2)

#Make clade table of mixing dates only
w_mixing <- bog_subset("..H", clade_table)
hits <- match(mixes, substr(colnames(w_mixing), start = 1, stop = 10))
hits <- hits[which(is.na(hits) == F)]
w_mixing <- w_mixing[,hits]


#Remove taxonomic level tags at the beginning of some names (present in Greengenes classified clades)
rownames(w_mixing) <- gsub("[kpcofg]__|\\[|\\]|Bacteria;", "", rownames(w_mixing))

#Keep only clades that are completely classified
splitnames<- strsplit(rownames(w_mixing), ";")
clades_only <- c()
for(i in 1:length(splitnames)){
  clades_only[i] <- splitnames[[i]][5]
}

w_mixing2 <- w_mixing[which(clades_only != "unclassified" & is.na(clades_only) == F),]
w_mixing2 <- reduce_names(w_mixing2)

#Subtract CoV of mixing dates from CoV of no mixing dates
Cov_no_mixing <- apply(hypo_clade_table2, 1, CoV)
Cov_w_mixing <- apply(w_mixing2, 1, CoV)
diff <- Cov_no_mixing - Cov_w_mixing
diff.top <- diff[match(names(hypo_clade_sums)[1:20], names(diff))]

#Make dataframe for plotting
plot.diff <- data.frame(names(diff.top), diff.top)
colnames(plot.diff) <- c("Clade", "Covariance")
plot.diff$Clade <- factor(plot.diff$Clade, levels=plot.diff$Clade)
ggplot(data=plot.diff, aes(x=Clade, y=Covariance)) + geom_bar(stat = "identity", fill = "grey", colour="black") + labs(x = NULL, y = "Difference in CoV") + theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 12, colour = "black"), axis.title = element_text(size = 12, vjust=2), axis.text.y = element_text(colour="black")) + scale_y_continuous(expand = c(0,0), limits = c(-2, 2)) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) + geom_hline(aes(yintercept=0))
```

####Figure S7. Difference in the coefficient of variance for the top 20 clades between hypolimnia (mixing dates and polymictic lake samples removed) and hypolimnia (mixing dates and polymictic lake samples included). A negative ratio of CoV indicates that a clade was more variable when mixed samples were included, and a positive ratio of CoV indicates that a clade was less variable when mixed samples were included. Clades of gamI and *Desulfobulbaceae* were less variable in mixed samples, while *Desulfobulbus* and *Deltaproteobacteria* were more variable in mixed samples.
***
```{r, PCoA:hypolimnia2007, echo=FALSE, warning=FALSE, message=FALSE}
#Make table of just samples from 2007 at the clade level
clade_table07 <- year_subset("07", clade_table)

#Make table of 2007 hypolimnion samples
all.hypo <- bog_subset("..H", clade_table07)

#Calculate distance matrix between samples using Bray-Curtis Similarity
distance <- vegdist(t(all.hypo), method = "bray")
#Label each point by its sampling location
lakes <- c("CB", "FB", "WS", "NS", "TB", "SS", "HK", "MA")
groups <- factor(substr(colnames(all.hypo), start = 1, stop = 2), levels = lakes)
#Run PCoA
pcoa <- betadisper(distance, groups)
#Extract scores for plotting
coord <- scores(pcoa)

#Make a vector to label each point by its mixing regime
regime <- c()
regime[which(groups == "CB" | groups == "FB" | groups == "WS")] <- "polymictic"
regime[which(groups == "NS" | groups == "TB" | groups == "SS")] <- "dimictic"
regime[which(groups == "HK" | groups == "MA")] <- "meromictic"

#Make dataframe for plotting
plot.pcoa <- data.frame(groups, coord$sites, regime)
colnames(plot.pcoa) <- c("Lake", "PCoA1", "PCoA2", "Regime")
ggplot(data=plot.pcoa, aes(x=PCoA1, y=PCoA2, colour=Lake, shape = Regime)) + geom_point(size=4) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) + theme(axis.text.x = element_text(hjust = 1, size = 12, colour = "black"), axis.title = element_text(size = 15, vjust=.7), axis.text.y = element_text(colour="black", size=12)) + scale_colour_brewer(palette = "Dark2")
```

####Figure S8a. Principle coordinates analysis of hypolimnia in all lakes. The color of each point indicates which lake it was sampled from, and the shape of each point indicates the mixing regime of that lake. Samples cluster by mixing regime, as well as by lake. Meromictic samples in particular are tightly clustered. 
***
```{r, PCoA:dispersion, echo=FALSE}
#Plot dispersion of each lakes' cluster.
boxplot(pcoa)
#Optional: run significance testing on clustering by lake. Takes a long time and will print straight to the markdown document.
#anosim(t(all.hypo), groups, permutations = 999, distance = "bray")
```

####Figure S8b. Distance from the centroid was calculated for each lake as a measure of dispersion. Lakes are arranged by order of increasing depth from left to right. Meromictic hypolimnia (MA and HK) are observed to have the lowest levels of dispersion, indicating a more stable community composition on long time scales. Polymictic hypolimnia (CB, FB, and WS) have increasing dispersion with depth. NS and TB, both dimictic have a wide range of dispersion, while dimictic SS appears more similar in dispersion to meromictic lakes.
***
```{r, PCoA:hypolimnia+years, echo=FALSE, warning=FALSE, message=FALSE}
#Plot PCoA of hypolimnia samples of each lake over all years of sampling available. The method is the same as above. This analysis is performed on lakes with multiple years of sampling: CB, TB, NS, SS, and MA. An example anosim() signficance test for clustering by year is included with TB: this takes a long time to run and prints directly to the markdown document. When this command was run for all lakes, all were significant at p=0.0001. 

TB <- bog_subset("TB.", clade_table)

distance <- vegdist(t(TB), method = "bray")
groups <- factor(substr(colnames(TB), start = 9, stop = 10), levels = c("05", "07", "08", "09"))

pcoa <- betadisper(distance, groups)
coord <- scores(pcoa)

layer <- factor(substr(colnames(TB), start=3, stop=3), levels = c("E", "H"))
plot.pcoa <- data.frame(groups, coord$sites, layer)
colnames(plot.pcoa) <- c("Year", "PCoA1", "PCoA2", "Layer")

ggplot(data=plot.pcoa, aes(x=PCoA1, y=PCoA2, colour=Year, shape=Layer)) + geom_point(size=3) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) + theme(axis.text.x = element_text(hjust = 1, size = 12, colour = "black"), axis.title = element_text(size = 15, vjust=.7), axis.text.y = element_text(colour="black", size=12)) + scale_colour_brewer(palette = "Dark2") + labs(title = "Trout Bog")
#anosim(t(TB), groups, permutations = 999, distance = "bray")
#TBH <- bog_subset("TBH", clade_table)
#group2 <- factor(substr(colnames(TBH), start = 9, stop = 10), levels = c("05", "07", "08", "09"))
#anosim(t(TBH), group2, permutations = 999, distance = "bray")
#TBE <- bog_subset("TBE", clade_table)
#group2 <- factor(substr(colnames(TBE), start = 9, stop = 10), levels = c("05", "07", "08", "09"))
#anosim(t(TBE), group2, permutations = 999, distance = "bray")
############
MA <- bog_subset("MA.", clade_table)

distance <- vegdist(t(MA), method = "bray")
groups <- factor(substr(colnames(MA), start = 9, stop = 10), levels = c("05", "07", "08", "09"))

pcoa <- betadisper(distance, groups)
coord <- scores(pcoa)

layer <- factor(substr(colnames(MA), start=3, stop=3), levels = c("E", "H"))
plot.pcoa <- data.frame(groups, coord$sites, layer)
colnames(plot.pcoa) <- c("Year", "PCoA1", "PCoA2", "Layer")

ggplot(data=plot.pcoa, aes(x=PCoA1, y=PCoA2, colour=Year, shape=Layer)) + geom_point(size=3) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) + theme(axis.text.x = element_text(hjust = 1, size = 12, colour = "black"), axis.title = element_text(size = 15, vjust=.7), axis.text.y = element_text(colour="black", size=12)) + scale_colour_brewer(palette = "Dark2")  + labs(title = "Mary Lake")
#anosim(t(MA), groups, permutations = 999, distance = "bray")
#MAH <- bog_subset("MAH", clade_table)
#group2 <- factor(substr(colnames(MAH), start = 9, stop = 10), levels = c("05", "07", "08", "09"))
#anosim(t(MAH), group2, permutations = 999, distance = "bray")
#MAE <- bog_subset("MAE", clade_table)
#group2 <- factor(substr(colnames(MAE), start = 9, stop = 10), levels = c("05", "07", "08", "09"))
#anosim(t(MAE), group2, permutations = 999, distance = "bray")
###########
CB <- bog_subset("CB.", clade_table)

groups <- factor(substr(colnames(CB), start = 9, stop = 10), levels = c("07", "09"))
layer <- factor(substr(colnames(CB[,which(is.na(groups) == F)]), start=3, stop=3), levels = c("E", "H"))
distance <- vegdist(t(CB[,which(is.na(groups) == F)]), method = "bray")
groups2 <- groups[which(is.na(groups) == F)]


pcoa <- betadisper(distance, groups2)
coord <- scores(pcoa)

plot.pcoa <- data.frame(groups2, coord$sites, layer)
colnames(plot.pcoa) <- c("Year", "PCoA1", "PCoA2", "Layer")

ggplot(data=plot.pcoa, aes(x=PCoA1, y=PCoA2, colour=Year, shape=Layer)) + geom_point(size=3) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) + theme(axis.text.x = element_text(hjust = 1, size = 12, colour = "black"), axis.title = element_text(size = 15, vjust=.7), axis.text.y = element_text(colour="black", size=12)) + scale_colour_brewer(palette = "Dark2")  + labs(title = "Crystal Bog")
#anosim(t(CB[,which(is.na(groups) == F)]), groups2, permutations = 999, distance = "bray")
#CBH <- bog_subset("CBH", clade_table)
#group2 <- factor(substr(colnames(CBH), start = 9, stop = 10), levels = c("07", "09"))
#anosim(t(CBH), group2, permutations = 999, distance = "bray")
#CBE <- bog_subset("CBE", clade_table)
#group2 <- factor(substr(colnames(CBE), start = 9, stop = 10), levels = c("07", "09"))
#anosim(t(CBE), group2, permutations = 999, distance = "bray")
############
NS <- bog_subset("NS.", clade_table)

groups <- factor(substr(colnames(NS), start = 9, stop = 10), levels = c("07", "08", "09"))
distance <- vegdist(t(NS[,which(is.na(groups) == F)]), method = "bray")
layer <- factor(substr(colnames(NS[,which(is.na(groups) == F)]), start=3, stop=3), levels = c("E", "H"))
groups2 <- groups[which(is.na(groups) == F)]

pcoa <- betadisper(distance, groups2)
coord <- scores(pcoa)


plot.pcoa <- data.frame(groups2, coord$sites, layer)
colnames(plot.pcoa) <- c("Year", "PCoA1", "PCoA2", "Layer")

ggplot(data=plot.pcoa, aes(x=PCoA1, y=PCoA2, colour=Year, shape=Layer)) + geom_point(size=3) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) + theme(axis.text.x = element_text(hjust = 1, size = 12, colour = "black"), axis.title = element_text(size = 15, vjust=.7), axis.text.y = element_text(colour="black", size=12)) + scale_colour_brewer(palette = "Dark2")  + labs(title = "North Sparkling Bog")
#anosim(t(NS[,which(is.na(groups) == F)]), groups2, permutations = 999, distance = "bray")
#NSH <- bog_subset("NSH", clade_table)
#group2 <- factor(substr(colnames(NSH), start = 9, stop = 10), levels = c("07", "08", "09"))
#NSH <- NSH[,which(is.na(group2) == F)]
#group2 <- group2[which(is.na(group2) == F)]
#anosim(t(NSH), group2 permutations = 999, distance = "bray")
#NSE <- bog_subset("NSE", clade_table)
#group2 <- factor(substr(colnames(NSE), start = 9, stop = 10), levels = c("07", "08", "09"))
#anosim(t(NSE), group2, permutations = 999, distance = "bray")
##########
SS <- bog_subset("SS.", clade_table)

distance <- vegdist(t(SS), method = "bray")
groups <- factor(substr(colnames(SS), start = 9, stop = 10), levels = c("07", "08", "09"))

pcoa <- betadisper(distance, groups)
coord <- scores(pcoa)

layer <- factor(substr(colnames(SS), start=3, stop=3), levels = c("E", "H"))
plot.pcoa <- data.frame(groups, coord$sites, layer)
colnames(plot.pcoa) <- c("Year", "PCoA1", "PCoA2", "Layer")

ggplot(data=plot.pcoa, aes(x=PCoA1, y=PCoA2, colour=Year, shape=Layer)) + geom_point(size=3) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) + theme(axis.text.x = element_text(hjust = 1, size = 12, colour = "black"), axis.title = element_text(size = 15, vjust=.7), axis.text.y = element_text(colour="black", size=12)) + scale_colour_brewer(palette = "Dark2")  + labs(title = "South Sparkling Bog")
#anosim(t(SS), groups, permutations = 999, distance = "bray")
#SSH <- bog_subset("SSH", clade_table)
#group2 <- factor(substr(colnames(SSH), start = 9, stop = 10), levels = c("05", "07", "08", "09"))
#anosim(t(SSH), group2, permutations = 999, distance = "bray")
#SSE <- bog_subset("SSE", clade_table)
#group2 <- factor(substr(colnames(SSE), start = 9, stop = 10), levels = c("05", "07", "08", "09"))
#anosim(t(SSE), group2, permutations = 999, distance = "bray")
```

####Figure S9. Principle coordinates analysis of epilimnion and hypolimnion samples over four years in all lakes. The shape of each point indicates whether it was collected from the epilimnion or hypolimnion, and the color indicates the year in which it was collected. Epilimnion samples look similar each year, while hypolimnion samples cluster by year. Mary Lake hypolimnion samples are highly similar each year, while Crystal Bog hypolimnion samples are more similar to epilimnion samples. However, all lakes and layers showed significant clustering by year (p less than or equal to 0.001) using an analysis of similarity test (anosim()).
***

Connectivity in all lakes over time?

```{r, Phenology, echo=FALSE, warning=FALSE, message=FALSE}
#Plot the phenological trend in similarity between dimictic and meromictic lakes. Trout Bog vs Mary Lake, 2007 is included in the main text of the paper. Below are phenologies between Mary Lake and other dimictic lakes and years.

metaNSH <- metadata[which(metalakes == "NSH" & metayears == "07"), c(1,2,4)]
metaNSH <- dcast(metaNSH, Sample_Name~Depth, fun.aggregate=mean)
NSHmixes <- metaNSH$Sample_Name[which(metaNSH$"0.5" - metaNSH$"4" < 1)]

bog1 <- "NSH"
bog2 <- "MAH"
table <- year_subset("07", clade_table)
query <- bog_subset(bog1, table)
database <- bog_subset(bog2, table)

mean.bc <- c()
for(i in 1:dim(query)[2]){
  bc.dis <- c()
  for(j in 1:dim(database)[2]){
    test <- cbind(query[,i], database[,j])
    bc.dis[j] <- 1- vegdist(t(test), method="bray")
  }
  mean.bc[i] <- mean(bc.dis) 
}

dates <- extract_date(colnames(query))
plot.data <- data.frame(dates, mean.bc)
colnames(plot.data) <- c("Dates", "BrayCurtis")
ggplot(data=plot.data, aes(x=Dates, y=BrayCurtis)) + geom_line(size=1.5) + geom_vline(xintercept = as.numeric(extract_date(NSHmixes)), linetype = "dashed", colour="black") + labs(y = "Bray-Curtis Similarity", title="North Sparkling vs Mary, 2007") + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) + theme(axis.text.x = element_text(vjust = 1.5, size = 12, colour = "black"), axis.title.x = element_text(size = 15), axis.text.y = element_text(colour="black", size=10), axis.title.y = element_text(size = 15, vjust=.5), axis.title.y = element_text(colour="black", size=15))

######

metaNSH <- metadata[which(metalakes == "NSH" & metayears == "09"), c(1,2,4)]
metaNSH <- dcast(metaNSH, Sample_Name~Depth, fun.aggregate=mean)
NSHmixes <- metaNSH$Sample_Name[which(metaNSH$"0.5" - metaNSH$"4" < 1)]

bog1 <- "NSH"
bog2 <- "MAH"
table <- year_subset("09", clade_table)
query <- bog_subset(bog1, table)
database <- bog_subset(bog2, table)

mean.bc <- c()
for(i in 1:dim(query)[2]){
  bc.dis <- c()
  for(j in 1:dim(database)[2]){
    test <- cbind(query[,i], database[,j])
    bc.dis[j] <- 1- vegdist(t(test), method="bray")
  }
  mean.bc[i] <- mean(bc.dis) 
}
#Only mix is in May when there is no sampling - not plotting for this graph
dates <- extract_date(colnames(query))
plot.data <- data.frame(dates, mean.bc)
  colnames(plot.data) <- c("Dates", "BrayCurtis")
  ggplot(data=plot.data, aes(x=Dates, y=BrayCurtis)) + geom_line(size=1.5) + labs(y = "Bray-Curtis Similarity", title="North Sparkling vs Mary, 2009") + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) + theme(axis.text.x = element_text(vjust = 1.5, size = 12, colour = "black"), axis.title.x = element_text(size = 15), axis.text.y = element_text(colour="black", size=10), axis.title.y = element_text(size = 15, vjust=.5), axis.title.y = element_text(colour="black", size=15))

#####
metaSSH <- metadata[which(metalakes == "SSH" & metayears == "08"), c(1,2,4)]
metaSSH <- dcast(metaSSH, Sample_Name~Depth, fun.aggregate=mean)
SSHmixes <- metaSSH$Sample_Name[which(metaSSH$"0.5" - metaSSH$"8" < 1)]

bog1 <- "SSH"
bog2 <- "MAH"
table <- year_subset("08", clade_table)
query <- bog_subset(bog1, table)
database <- bog_subset(bog2, table)

mean.bc <- c()
for(i in 1:dim(query)[2]){
  bc.dis <- c()
  for(j in 1:dim(database)[2]){
    test <- cbind(query[,i], database[,j])
    bc.dis[j] <- 1- vegdist(t(test), method="bray")
  }
  mean.bc[i] <- mean(bc.dis) 
}

dates <- extract_date(colnames(query))
plot.data <- data.frame(dates, mean.bc)
  colnames(plot.data) <- c("Dates", "BrayCurtis")
  ggplot(data=plot.data, aes(x=Dates, y=BrayCurtis)) + geom_line(size=1.5) + geom_vline(xintercept = as.numeric(extract_date(SSHmixes)), linetype = "dashed", colour="black") + labs(y = "Bray-Curtis Similarity", title="South Sparkling vs Mary, 2008") + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) + theme(axis.text.x = element_text(vjust = 1.5, size = 12, colour = "black"), axis.title.x = element_text(size = 15), axis.text.y = element_text(colour="black", size=10), axis.title.y = element_text(size = 15, vjust=.5), axis.title.y = element_text(colour="black", size=15))

#####
metaSSH <- metadata[which(metalakes == "SSH" & metayears == "07"), c(1,2,4)]
metaSSH <- dcast(metaSSH, Sample_Name~Depth, fun.aggregate=mean)
SSHmixes <- metaSSH$Sample_Name[which(metaSSH$"0.5" - metaSSH$"8" < 1)]

bog1 <- "SSH"
bog2 <- "MAH"
table <- year_subset("07", clade_table)
query <- bog_subset(bog1, table)
database <- bog_subset(bog2, table)

mean.bc <- c()
for(i in 1:dim(query)[2]){
  bc.dis <- c()
  for(j in 1:dim(database)[2]){
    test <- cbind(query[,i], database[,j])
    bc.dis[j] <- 1- vegdist(t(test), method="bray")
  }
  mean.bc[i] <- mean(bc.dis) 
}

dates <- extract_date(colnames(query))
plot.data <- data.frame(dates, mean.bc)
  colnames(plot.data) <- c("Dates", "BrayCurtis")
  ggplot(data=plot.data, aes(x=Dates, y=BrayCurtis)) + geom_line(size=1.5) + geom_vline(xintercept = as.numeric(extract_date(SSHmixes)), linetype = "dashed", colour="black") + labs(y = "Bray-Curtis Similarity", title="South Sparkling vs Mary, 2007") + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) + theme(axis.text.x = element_text(vjust = 1.5, size = 12, colour = "black"), axis.title.x = element_text(size = 15), axis.text.y = element_text(colour="black", size=10), axis.title.y = element_text(size = 15, vjust=.5), axis.title.y = element_text(colour="black", size=15))

#####
metaSSH <- metadata[which(metalakes == "SSH" & metayears == "09"), c(1,2,4)]
metaSSH <- dcast(metaSSH, Sample_Name~Depth, fun.aggregate=mean)
SSHmixes <- metaSSH$Sample_Name[which(metaSSH$"0.5" - metaSSH$"8" < 1)]

bog1 <- "SSH"
bog2 <- "MAH"
table <- year_subset("09", clade_table)
query <- bog_subset(bog1, table)
database <- bog_subset(bog2, table)

mean.bc <- c()
for(i in 1:dim(query)[2]){
  bc.dis <- c()
  for(j in 1:dim(database)[2]){
    test <- cbind(query[,i], database[,j])
    bc.dis[j] <- 1- vegdist(t(test), method="bray")
  }
  mean.bc[i] <- mean(bc.dis) 
}

dates <- extract_date(colnames(query))
plot.data <- data.frame(dates, mean.bc)
  colnames(plot.data) <- c("Dates", "BrayCurtis")
  ggplot(data=plot.data, aes(x=Dates, y=BrayCurtis)) + geom_line(size=1.5) + geom_vline(xintercept = as.numeric(extract_date(SSHmixes)), linetype = "dashed", colour="black") + labs(y = "Bray-Curtis Similarity", title="South Sparkling vs Mary, 2009") + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) + theme(axis.text.x = element_text(vjust = 1.5, size = 12, colour = "black"), axis.title.x = element_text(size = 15), axis.text.y = element_text(colour="black", size=10), axis.title.y = element_text(size = 15, vjust=.5), axis.title.y = element_text(colour="black", size=15))

######
metaTBH <- metadata[which(metalakes == "TBH" & metayears == "08"), c(1,2,4)]
metaTBH <- dcast(metaTBH, Sample_Name~Depth, fun.aggregate=mean)
TBHmixes <- metaTBH$Sample_Name[which(metaTBH$"0.5" - metaTBH$"6" < 1)]

bog1 <- "TBH"
bog2 <- "MAH"
table <- year_subset("08", clade_table)
query <- bog_subset(bog1, table)
database <- bog_subset(bog2, table)

mean.bc <- c()
for(i in 1:dim(query)[2]){
  bc.dis <- c()
  for(j in 1:dim(database)[2]){
    test <- cbind(query[,i], database[,j])
    bc.dis[j] <- 1- vegdist(t(test), method="bray")
  }
  mean.bc[i] <- mean(bc.dis) 
}

dates <- extract_date(colnames(query))
plot.data <- data.frame(dates, mean.bc)
  colnames(plot.data) <- c("Dates", "BrayCurtis")
  ggplot(data=plot.data, aes(x=Dates, y=BrayCurtis)) + geom_line(size=1.5) + geom_vline(xintercept = as.numeric(extract_date(TBHmixes)), linetype = "dashed", colour="black") + labs(y = "Bray-Curtis Similarity", title="Trout Bog vs Mary, 2008") + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) + theme(axis.text.x = element_text(vjust = 1.5, size = 12, colour = "black"), axis.title.x = element_text(size = 15), axis.text.y = element_text(colour="black", size=10), axis.title.y = element_text(size = 15, vjust=.5), axis.title.y = element_text(colour="black", size=15))

######
metaTBH <- metadata[which(metalakes == "TBH" & metayears == "05"), c(1,2,4)]
metaTBH <- dcast(metaTBH, Sample_Name~Depth, fun.aggregate=mean)
TBHmixes <- metaTBH$Sample_Name[which(metaTBH$"0.5" - metaTBH$"6" < 1)]

bog1 <- "TBH"
bog2 <- "MAH"
table <- year_subset("05", clade_table)
query <- bog_subset(bog1, table)
database <- bog_subset(bog2, table)

mean.bc <- c()
for(i in 1:dim(query)[2]){
  bc.dis <- c()
  for(j in 1:dim(database)[2]){
    test <- cbind(query[,i], database[,j])
    bc.dis[j] <- 1- vegdist(t(test), method="bray")
  }
  mean.bc[i] <- mean(bc.dis) 
}

dates <- extract_date(colnames(query))
plot.data <- data.frame(dates, mean.bc)
  colnames(plot.data) <- c("Dates", "BrayCurtis")
  ggplot(data=plot.data, aes(x=Dates, y=BrayCurtis)) + geom_line(size=1.5) + geom_vline(xintercept = as.numeric(extract_date(TBHmixes)), linetype = "dashed", colour="black") + labs(y = "Bray-Curtis Similarity", title="Trout Bog vs Mary, 2005") + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) + theme(axis.text.x = element_text(vjust = 1.5, size = 12, colour = "black"), axis.title.x = element_text(size = 15), axis.text.y = element_text(colour="black", size=10), axis.title.y = element_text(size = 15, vjust=.5), axis.title.y = element_text(colour="black", size=15))

######
metaTBH <- metadata[which(metalakes == "TBH" & metayears == "09"), c(1,2,4)]
metaTBH <- dcast(metaTBH, Sample_Name~Depth, fun.aggregate=mean)
TBHmixes <- metaTBH$Sample_Name[which(metaTBH$"0.5" - metaTBH$"6" < 1)]

bog1 <- "TBH"
bog2 <- "MAH"
table <- year_subset("09", clade_table)
query <- bog_subset(bog1, table)
database <- bog_subset(bog2, table)

mean.bc <- c()
for(i in 1:dim(query)[2]){
  bc.dis <- c()
  for(j in 1:dim(database)[2]){
    test <- cbind(query[,i], database[,j])
    bc.dis[j] <- 1- vegdist(t(test), method="bray")
  }
  mean.bc[i] <- mean(bc.dis) 
}

dates <- extract_date(colnames(query))
plot.data <- data.frame(dates, mean.bc)
  colnames(plot.data) <- c("Dates", "BrayCurtis")
  ggplot(data=plot.data, aes(x=Dates, y=BrayCurtis)) + geom_line(size=1.5) + labs(y = "Bray-Curtis Similarity", title="Trout Bog vs Mary, 2009") + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) + theme(axis.text.x = element_text(vjust = 1.5, size = 12, colour = "black"), axis.title.x = element_text(size = 15), axis.text.y = element_text(colour="black", size=10), axis.title.y = element_text(size = 15, vjust=.5), axis.title.y = element_text(colour="black", size=15))

######
metaCBH <- metadata[which(metalakes == "CBH" & metayears == "07"), c(1,2,4)]
metaCBH <- dcast(metaCBH, Sample_Name~Depth, fun.aggregate=mean)
CBHmixes <- metaCBH$Sample_Name[which(metaCBH$"0.5" - metaCBH$"2" < 1)]

bog1 <- "CBH"
bog2 <- "MAH"
table <- year_subset("07", clade_table)
query <- bog_subset(bog1, table)
database <- bog_subset(bog2, table)

mean.bc <- c()
for(i in 1:dim(query)[2]){
  bc.dis <- c()
  for(j in 1:dim(database)[2]){
    test <- cbind(query[,i], database[,j])
    bc.dis[j] <- 1- vegdist(t(test), method="bray")
  }
  mean.bc[i] <- mean(bc.dis) 
}

dates <- extract_date(colnames(query))
plot.data <- data.frame(dates, mean.bc)
  colnames(plot.data) <- c("Dates", "BrayCurtis")
  ggplot(data=plot.data, aes(x=Dates, y=BrayCurtis)) + geom_line(size=1.5) + labs(y = "Bray-Curtis Similarity", title="Crystal Bog vs Mary, 2007") + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) + theme(axis.text.x = element_text(vjust = 1.5, size = 12, colour = "black"), axis.title.x = element_text(size = 15), axis.text.y = element_text(colour="black", size=10), axis.title.y = element_text(size = 15, vjust=.5), axis.title.y = element_text(colour="black", size=15))

#####
metaCBH <- metadata[which(metalakes == "CBH" & metayears == "09"), c(1,2,4)]
metaCBH <- dcast(metaCBH, Sample_Name~Depth, fun.aggregate=mean)
CBHmixes <- metaCBH$Sample_Name[which(metaCBH$"0.5" - metaCBH$"2" < 1)]

bog1 <- "CBH"
bog2 <- "MAH"
table <- year_subset("09", clade_table)
query <- bog_subset(bog1, table)
database <- bog_subset(bog2, table)

mean.bc <- c()
for(i in 1:dim(query)[2]){
  bc.dis <- c()
  for(j in 1:dim(database)[2]){
    test <- cbind(query[,i], database[,j])
    bc.dis[j] <- 1- vegdist(t(test), method="bray")
  }
  mean.bc[i] <- mean(bc.dis) 
}

dates <- extract_date(colnames(query))
plot.data <- data.frame(dates, mean.bc)
  colnames(plot.data) <- c("Dates", "BrayCurtis")
  ggplot(data=plot.data, aes(x=Dates, y=BrayCurtis)) + geom_line(size=1.5) + labs(y = "Bray-Curtis Similarity", title="Crystal Bog vs Mary, 2009") + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) + theme(axis.text.x = element_text(vjust = 1.5, size = 12, colour = "black"), axis.title.x = element_text(size = 15), axis.text.y = element_text(colour="black", size=10), axis.title.y = element_text(size = 15, vjust=.5), axis.title.y = element_text(colour="black", size=15))

#####
metaFBH <- metadata[which(metalakes == "FBH" & metayears == "07"), c(1,2,4)]
metaFBH <- dcast(metaFBH, Sample_Name~Depth, fun.aggregate=mean)
FBHmixes <- metaFBH$Sample_Name[which(metaFBH$"0.5" - metaFBH$"2" < 1)]

bog1 <- "FBH"
bog2 <- "MAH"
table <- year_subset("07", clade_table)
query <- bog_subset(bog1, table)
database <- bog_subset(bog2, table)

mean.bc <- c()
for(i in 1:dim(query)[2]){
  bc.dis <- c()
  for(j in 1:dim(database)[2]){
    test <- cbind(query[,i], database[,j])
    bc.dis[j] <- 1- vegdist(t(test), method="bray")
  }
  mean.bc[i] <- mean(bc.dis) 
}

dates <- extract_date(colnames(query))
plot.data <- data.frame(dates, mean.bc)
  colnames(plot.data) <- c("Dates", "BrayCurtis")
  ggplot(data=plot.data, aes(x=Dates, y=BrayCurtis)) + geom_line(size=1.5) + labs(y = "Bray-Curtis Similarity", title="Forestry Bog vs Mary, 2007") + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) + theme(axis.text.x = element_text(vjust = 1.5, size = 12, colour = "black"), axis.title.x = element_text(size = 15), axis.text.y = element_text(colour="black", size=10), axis.title.y = element_text(size = 15, vjust=.5), axis.title.y = element_text(colour="black", size=15))

#####
metaWSH <- metadata[which(metalakes == "WSH" & metayears == "07"), c(1,2,4)]
metaWSH <- dcast(metaWSH, Sample_Name~Depth, fun.aggregate=mean)
WSHmixes <- metaWSH$Sample_Name[which(metaWSH$"0.5" - metaWSH$"4" < 1)]

bog1 <- "WSH"
bog2 <- "MAH"
table <- year_subset("07", clade_table)
query <- bog_subset(bog1, table)
database <- bog_subset(bog2, table)

mean.bc <- c()
for(i in 1:dim(query)[2]){
  bc.dis <- c()
  for(j in 1:dim(database)[2]){
    test <- cbind(query[,i], database[,j])
    bc.dis[j] <- 1- vegdist(t(test), method="bray")
  }
  mean.bc[i] <- mean(bc.dis) 
}

dates <- extract_date(colnames(query))
plot.data <- data.frame(dates, mean.bc)
  colnames(plot.data) <- c("Dates", "BrayCurtis")
  ggplot(data=plot.data, aes(x=Dates, y=BrayCurtis)) + geom_line(size=1.5) + labs(y = "Bray-Curtis Similarity", title="West Sparkling Bog vs Mary, 2007") + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) + theme(axis.text.x = element_text(vjust = 1.5, size = 12, colour = "black"), axis.title.x = element_text(size = 15), axis.text.y = element_text(colour="black", size=10), axis.title.y = element_text(size = 15, vjust=.5), axis.title.y = element_text(colour="black", size=15))

#####
```

####Figure S10. Bray-Curtis Similarity between each dimictic bog lake and Mary Lake . Each point represents the average similarity between the dimictic or polymictic lake sample taken on that date and every Mary lake sample taken that year. Dashed lines represent days when the water column was mixed, defined as less than one degree difference in temperature between 0.5 m and the maximum sampling depth.   
***