---
title: "Supplemental Figures"
author: "Alexandra Linz"
date: "Thursday, January 28, 2016"
output: pdf_document
---
###Supplemental figures accompanying the paper "Seasonal mixing frequency and historical mixing regime determine hypolimnion bacterial community dynamics in bog lakes"

```{r, Setting_up_the_environoment, echo=F, warning=FALSE, message=FALSE}
# Input your path to the North_Temperate_Lakes-Microbial_Observatory/Figures folder in the GitHub repo, or whereever you would like the figures to be saved
path2repo <- "C:/Users/amlinz16/Desktop/North_Temperate_Lakes-Microbial_Observatory/Figures/"

# Load packages
library(OTUtable)       # Contains data and functions for analysis of NTL-MO OTU table
library(vegan)          # Used for Bray-Curtis
library(ggplot2)        # Used for plotting
library(reshape2)       # Used to format metadata
library(ggdendro)       # Used to plot hierarchical clustering trees
library(exactRankTests) # Used to test evenness across lakes
# Load data from OTUtable
data(otu_table)
data(taxonomy)
data(metadata)

# Generate tables at the clade and phylum level
clade_table <- combine_otus("Clade", otu_table, taxonomy)
phylum_table <- combine_otus("Phylum", otu_table, taxonomy)



```

```{r, Evenness between bogs, echo=FALSE, warning=FALSE, message=FALSE}
# List lake categories
lakes <- c("CB", "FB", "WS", "NS", "TB", "SS", "HK", "MA")

# Split epilimnion and hypolimnion into separate tables
epilimnia <- bog_subset("..E", otu_table)
hypolimnia <- bog_subset("..H", otu_table)

# Calculate observed richness
epi.evenness <- apply(epilimnia, 2, pielou)
hypo.evenness <- apply(hypolimnia, 2, pielou)

# Extract sampling location from sample names
epi.lakes <- substr(names(epi.evenness), start = 1, stop = 2)
hypo.lakes <- substr(names(hypo.evenness), start = 1, stop = 2)

#Make dataframe for plotting
epi.data <- data.frame(epi.lakes, epi.evenness)
hypo.data <- data.frame(hypo.lakes, hypo.evenness)
epi.data$epi.lakes <- ordered(epi.data$epi.lakes, levels = lakes)
hypo.data$hypo.lakes <- ordered(hypo.data$hypo.lakes, levels = lakes)

ggplot(data = epi.data, aes(y = epi.evenness, x = epi.lakes, fill = epi.lakes)) + geom_boxplot() + labs(y = "Observed Evenness", x = NULL, title = "Epilimnion") + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) + theme(axis.text.x = element_text(hjust = 1, size = 16, colour = "black"), axis.title = element_text(size = 18, vjust = 2), axis.text.y = element_text(colour = "black", size = 14), legend.position = "none") + scale_fill_brewer(palette = "Set3")

ggplot(data = hypo.data, aes(y = hypo.evenness, x = hypo.lakes, fill = hypo.lakes)) + geom_boxplot() + labs(y = "Observed Evenness", x = NULL, title = "Hypolimnion") + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) + theme(axis.text.x = element_text(hjust = 1, size = 16, colour = "black"), axis.title = element_text(size = 18, vjust = 2), axis.text.y = element_text(colour = "black", size = 14), legend.position = "none") + scale_fill_brewer(palette = "Set3") 


#Check significance using Wilcoxon signed-rank test

lakeids <- levels(epi.data$epi.lakes)
Lake1 <- character(0)
Lake2 <- character(0)
Test <- character(0)
pvalue <- numeric(0)
Interpretation <- character(0)

for (id in lakeids) { 
  x = subset(epi.data, epi.data$epi.lakes == id)
  for (lake in lakeids) {
    y = subset(epi.data, epi.data$epi.lakes == lake)
    result <- wilcox.exact(x$epi.evenness,y$epi.evenness, alternative= "two.sided", conf.level = 0.95)
    Lake1 <- c(Lake1, id)
    Lake2 <- c(Lake2, lake)
    Test <- c(Test, result$method)
    pvalue <- c(pvalue, result$p.value)
    if (result$p.value < 0.05) {
      Interpretation <- c(Interpretation, "Reject Null (difference in median detected)")
    } else {
      Interpretation <- c(Interpretation, "Accept Null (difference in median undetected)")
    }  
  }
}

Wilcoxon.epi <- data.frame(Lake1, Lake2, Test, pvalue, Interpretation)
#print(Wilcoxon.epi)

lakeids <- levels(hypo.data$hypo.lakes)
Lake1 <- character(0)
Lake2 <- character(0)
Test <- character(0)
pvalue <- numeric(0)
Interpretation <- character(0)

for (id in lakeids) { 
  x = subset(hypo.data, hypo.data$hypo.lakes == id)
  for (lake in lakeids) {
    y = subset(hypo.data, hypo.data$hypo.lakes == lake)
    result <- wilcox.exact(x$hypo.evenness,y$hypo.evenness, alternative= "two.sided", conf.level = 0.95)
    Lake1 <- c(Lake1, id)
    Lake2 <- c(Lake2, lake)
    Test <- c(Test, result$method)
    pvalue <- c(pvalue, result$p.value)
    if (result$p.value < 0.05) {
      Interpretation <- c(Interpretation, "Reject Null (difference in median detected)")
    } else {
      Interpretation <- c(Interpretation, "Accept Null (difference in median undetected)")
    }  
  }
}

Wilcoxon.hypo <- data.frame(Lake1, Lake2, Test, pvalue, Interpretation)
#print(Wilcoxon.hypo)

```

####Figure S1. Evenness by lakes. Pielou's Evenness metric was used to measure the evenness of each sample and grouped by lake. While most pairs of lakes are significantly different in richness in both layers, no trend of increasing or decreasing evenness with depth is observed. However, the evenness in the hypolimnia of both meromictic lakes is significantly greater than all other lakes in this study.
***
```{r, Evenness over time, echo=FALSE, warning=FALSE, message=FALSE}

# Identify mixing dates (less than 1 degree of temperature difference between 0.5 meters and maximum sampling depth)
metalakes <- substr(metadata$Sample_Name, start = 1, stop = 3)
metayears <- substr(metadata$Sample_Name, start = 9, stop = 10)
metaTBH <- metadata[which(metalakes == "TBH" & metayears == "07"), c(1,2,4)]
metaTBH <- dcast(metaTBH, Sample_Name~Depth, fun.aggregate=mean)
TBHmixes <- extract_date(metaTBH$Sample_Name[which(metaTBH$"0.5" - metaTBH$"7" < 1)])

# Make dataset of Trout Bog hypolimion samples from 2007
hypo <- bog_subset(paste("TBH", sep = ""), otu_table)
hypo <- year_subset("07", hypo)
#Calculate observed richness
hypo.even <- apply(hypo, 2, pielou)
#Extract sampling date from sample names
hypo.date <- extract_date(colnames(hypo))
#Remove January samples - large gap distracts in plot, and winter samples are not considered in this study
hypo.even <- hypo.even[c(1:32, 35:80)]
hypo.date <- hypo.date[c(1:32, 35:80)]

# Make dataframe for plotting
TB_evenness <- data.frame(hypo.date, hypo.even)
colnames(TB_evenness) <- c("date", "evenness")

ggplot() + geom_line(data = TB_evenness, aes(x = date, y=evenness), size = 1) + labs(title = "Trout Bog", x = NULL, y = "Pielou's Evenness") + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) + theme(axis.text.x = element_text(hjust = 1, size = 8, colour = "black"), axis.title = element_text(size = 12, vjust = 2), axis.text.y = element_text(colour = "black", size = 8), plot.title = element_text(size = 12, vjust = 2)) + geom_point(data = TB_evenness[match(TBHmixes, TB_evenness$date), ], aes(x = date, y = evenness), size = 5, colour = "red") 

# Repeat with North Sparkling, 2008
metaNSH <- metadata[which(metalakes == "NSH" & metayears == "08"), c(1,2,4)]
metaNSH <- dcast(metaNSH, Sample_Name~Depth, fun.aggregate=mean)
NSHmixes <- extract_date(metaNSH$Sample_Name[which(metaNSH$"0.5" - metaNSH$"4" < 1)])

#Make dataset of Trout Bog hypolimion samples from 2008
hypo <- bog_subset(paste("NSH", sep = ""), otu_table)
hypo <- year_subset("08", hypo)
#Calculate observed richness
hypo.even <- apply(hypo, 2, pielou)
#Extract sampling date from sample names
hypo.date <- extract_date(colnames(hypo))

#Make dataframe for plotting
NS_evenness <- data.frame(hypo.date, hypo.even)
colnames(NS_evenness) <- c("date", "evenness")

ggplot() + geom_line(data = NS_evenness, aes(x = date, y=evenness), size = 1) + labs(title = "North Sparkling Bog", x = NULL, y = "Pielou's Evenness") + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) + theme(axis.text.x = element_text(hjust = 1, size = 8, colour = "black"), axis.title = element_text(size = 12, vjust = 2), axis.text.y = element_text(colour = "black", size = 8), plot.title = element_text(size = 12, vjust = 2)) + geom_point(data = NS_evenness[match(NSHmixes, NS_evenness$date), ], aes(x = date, y = evenness), size = 5, colour = "red") 

```

####Figure S2. Evenness over time. Evenness decreases during the Trout Bog 2007 fall mixing event, as is observed for richness. During the North Sparkling Bog artificial mixing event in 2008, evenness decreases but quickly recovers, with no drop in evenness during the fall mixing event for that year. 
***

```{r, Phylum_abundance, echo=FALSE, warning=FALSE, message=FALSE}
# Keep only phyla labels in phylum_table
phylum_table <- reduce_names(phylum_table)

# Sum total observations of each phylum in the full dataset and sort by abundance
totals <- rowSums(phylum_table)
totals <- sort(totals)

# Remove the "p__" phylum designation from phylum names. This looks better for plotting.
get.names <- strsplit(names(totals), "p__")
phyla.names <- c()
for(i in 1:length(get.names)){
  phyla.names[i] <- get.names[[i]][2]
}
phyla.names[which(is.na(phyla.names) == T)] <- "unclassified"

# Set up a dataframe for plotting in ggplot2. Set the phyla.names to factors, with levels in order of abundance.
phylum_totals <- data.frame(phyla.names, totals)
phylum_totals$phyla.names <- factor(phylum_totals$phyla.names, levels = phylum_totals$phyla.names[order(phylum_totals$totals)])

# Plot as a bar graph
ggplot(data=phylum_totals, aes(x = phyla.names, y = totals)) + geom_bar(stat = "identity", fill = "grey", colour="black") + labs(x = NULL, y = "Log of Observed Reads") + theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 12, colour = "black"), axis.title = element_text(size = 15, vjust=1.2), plot.title = element_text(size = 20), axis.text.y = element_text(colour="black")) + scale_y_log10(expand = c(0,0)) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black"))

```

####Figure S3. Forty phyla were identified over the entire NTL-MO time series. See Table 1 for lake abbrevations; the third letter indicates the sampling depth as "E" for epilimnion and "H" for hypolimnion. A) The most abundant phyla were Proteobacteria, Actinobacteria, Bacteriodetes, and Verrucomicrobia. Reads that could not be classified into a phylum were the fourth most abundant category. Candidate phyla members such as OP3 and SR1 are present in this dataset. 
***

```{r, Phylum_barchart, echo=FALSE, message=FALSE, warning=FALSE}
# Specify the sampled layers
layer <- c("CBE", "FBE", "WSE", "NSE", "TBE", "SSE", "HKE", "MAE", "CBH", "FBH", "WSH", "NSH", "TBH",  "SSH", "HKH", "MAH")

# Set up a dataframe with the totals of each phylum for each sampling site
layer.phyla <- rep(NA, dim(phylum_table)[1])

for(i in 1:length(layer)){
  dataset <- bog_subset(layer[i], phylum_table)
  layer.phyla <- cbind(layer.phyla, rowSums(dataset)) 
}

layer.phyla <- layer.phyla[,2:dim(layer.phyla)[2]]
colnames(layer.phyla) <- layer

# Combine low abundance phyla into a single category called "other"
abun <- layer.phyla[which(rowSums(layer.phyla) >= 10000),]
other <- layer.phyla[which(rowSums(layer.phyla) < 10000),]
new.layer <- rbind(abun, colSums(other))

# Shorten up those names again and remove extraneous rownames
get.names <- strsplit(rownames(new.layer), "p__")
phyla.names <- c()
for(i in 1:length(get.names)){
  phyla.names[i] <- get.names[[i]][2]
}
phyla.names[7] <- "unclassifed"
phyla.names[12] <- "other"
phyla.names <- factor(phyla.names, levels = rev(phyla.names))
rownames(new.layer) <- NULL

# Convert data into a long format dataframe for use in ggplot
phyla_by_bog <- data.frame(phyla.names, new.layer)
phyla_by_bog2 <- melt(phyla_by_bog)

# Create color palette that can handle the large number of categories
pal2 = c("#005682", "#edfb48", "#a1a100", "#626262", "#008141", "#008282", "#00d5f2", "#f2a400", "#209401", "#929292", "#3885e7", "#ff8400")

# Plot data as a stacked barplot
ggplot(data=phyla_by_bog2, aes(x=variable, y=value, fill=phyla.names)) + geom_bar(stat="identity", position = "fill") + labs(x = NULL, y = "Proportion of Observed Reads") + theme(axis.text.x = element_text(size = 12, angle = 90, color="black"), axis.text.y = element_text(size=14, color="black"), axis.title = element_text(size = 15, vjust=2), legend.title = element_blank(), legend.text = element_text(size = 16)) + scale_fill_manual(values=rev(pal2)) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black"))  + scale_y_continuous(expand = c(0,0)) 

```

####Figure S4. Phylum-level community differences were compared between layers and lakes. Epilimnion samples were relatively similar across all lakes, with the exception of the presence of Planctomyces in HKE and MAE, the epilimnia of the two meromictic lakes included in this study. Hypolimnion samples showed a trend of decreasing Actinobacteria with total lake depth, while the numbers of unclassified bacteria and other phyla (those comprising less than 8,000 reads in the dataset) increase with total lake depth. The sampled communities in hypolimnia of dimictic lakes (NSH, TBH, and SSH) contained higher levels of Acidobacteria than the other categories.
***

```{r, Rank_abundance:epilimnia, echo=FALSE, warning=FALSE, message=FALSE}
# Keep only epilimnion samples
epi_clade_table <- bog_subset("..E", clade_table)

# Remove taxonomic level tags at the beginning of some names (present in Greengenes classified clades)
rownames(epi_clade_table) <- make.unique(gsub("[kpcofg]__|\\[|\\]|Bacteria;", "", rownames(epi_clade_table)))

# Keep only clades that are completely classified
splitnames<- strsplit(rownames(epi_clade_table), ";")
clades_only <- c()
for(i in 1:length(splitnames)){
  clades_only[i] <- splitnames[[i]][5]
}

epi_clade_table2 <- epi_clade_table[which(clades_only != "unclassified" & is.na(clades_only) == F),]
epi_clade_table2 <- reduce_names(epi_clade_table2)
# Sum total abundances of epilimnion clades
epi_clade_sums <- sort(rowSums(epi_clade_table2), decreasing = T)

# Make dataframe of the 20 most abundant epilimnion clades for plotting
plot_epi_clades <- data.frame(names(epi_clade_sums)[1:20], epi_clade_sums[1:20])
colnames(plot_epi_clades) <- c("Clade", "Abundance")
plot_epi_clades$Clade <- factor(plot_epi_clades$Clade, levels=plot_epi_clades$Clade)
ggplot(data = plot_epi_clades, aes(x = Clade, y = Abundance)) + geom_bar(stat = "identity", fill = "grey", colour = "black") + labs(x = NULL, y = "Observed Reads in Epilimnia") + theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 12, colour = "black"), axis.title = element_text(size = 12, vjust=2), axis.text.y = element_text(colour="black")) + scale_y_continuous(expand = c(0,0)) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black"))

# Repeat the same analysis as above on hypolimnion samples, but remove samples where the water column was mixed, as well as all polymictic lake samples.
# I use the metadata file to find samples from mixed days. I define a mixed day as less than 1 degree of temperature difference between 0.5 meters and the maximum sampling depth.

metadata$Sample_Name <- as.character(metadata$Sample_Name)
metalakes <- substr(metadata$Sample_Name, start=1, stop=3)
metayears <- substr(metadata$Sample_Name, start=9, stop=10)

# Identify mixed dates from dimictic lakes
metaTBH <- metadata[which(metalakes == "TBH" & metayears == "07"), c(1,2,4)]
metaTBH <- dcast(metaTBH, Sample_Name~Depth, fun.aggregate=mean)
TBHmixes <- metaTBH$Sample_Name[which(metaTBH$"0.5" - metaTBH$"7" < 1)]

metaNSH <- metadata[which(metalakes == "NSH" & metayears == "07"), c(1,2,4)]
metaNSH <- dcast(metaNSH, Sample_Name~Depth, fun.aggregate=mean)
NSHmixes <- metaNSH$Sample_Name[which(metaNSH$"0.5" - metaNSH$"4" < 1)]

metaSSH <- metadata[which(metalakes == "SSH" & metayears == "07"), c(1,2,4)]
metaSSH <- dcast(metaSSH, Sample_Name~Depth, fun.aggregate=mean)
SSHmixes <- metaSSH$Sample_Name[which(metaSSH$"0.5" - metaSSH$"8" < 1)]

mixes <- c(TBHmixes, NSHmixes, SSHmixes)

# Remove mixed dates from table of dimictic and meromictic hypolimnion samples
hypo_clade_table <- bog_subset("TBH|SSH|NSH|MAH|HKH", clade_table)
hits <- match(mixes, substr(colnames(hypo_clade_table), start = 1, stop = 10))
hypo_clade_table <- hypo_clade_table[, -hits]

# Remove taxonomic level tags at the beginning of some names (present in Greengenes classified clades)
rownames(hypo_clade_table) <- gsub("[kpcofg]__|\\[|\\]|Bacteria;", "", rownames(hypo_clade_table))

# Keep only clades that are completely classified
splitnames<- strsplit(rownames(hypo_clade_table), ";")
clades_only <- c()
for(i in 1:length(splitnames)){
  clades_only[i] <- splitnames[[i]][5]
}

hypo_clade_table2 <- hypo_clade_table[which(clades_only != "unclassified" & is.na(clades_only) == F),]
hypo_clade_table2 <- reduce_names(hypo_clade_table2)

# Plot rank abundance curve for 20 most abundant hypolimnion clades
hypo_clade_sums <- sort(rowSums(hypo_clade_table2), decreasing = T)
plot_hypo_clades <- data.frame(names(hypo_clade_sums)[1:20], hypo_clade_sums[1:20])
colnames(plot_hypo_clades) <- c("Clade", "Abundance")
plot_hypo_clades$Clade <- factor(plot_hypo_clades$Clade, levels=plot_hypo_clades$Clade)
ggplot(data = plot_hypo_clades, aes(x = Clade, y = Abundance)) + geom_bar(stat = "identity", fill = "grey", colour="black") + labs(x = NULL, y = "Observed Reads in Hypolimnia") + theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 12, colour = "black"), axis.title = element_text(size = 12, vjust = 2), axis.text.y = element_text(colour = "black")) + scale_y_continuous(expand = c(0,0)) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black"))
```

####Figure S5. Rank abundance of the 20 most abundant clades in epilimnia samples. Total reads in all epilimnion samples have been summed for each clade-level group. Names represent the last classified level of taxonomic information for that clade; for example, the designation "Bacteria" indicates that clades unclassified at the phylum level are the fifth most abundant group in these samples. The most abundant groups in epilimnia are Pnec, acI-B, betI-A, and bacI-A. 
***

```{r, Connectivity between bogs, echo=FALSE}
all.network <- read.table(file = "C:/Users/amlinz16/Desktop/North_Temperate_Lakes-Microbial_Observatory/Network_analysis/allsamples_network_28Jan16.txt", header = T)
TBH <- bog_subset("TBH", otu_table)
NSH <- bog_subset("NSH", otu_table)
MAH <- bog_subset("MAH", otu_table)
SSH <- bog_subset("SSH", otu_table)
CBH <- bog_subset("CBH", otu_table)

TBH.edges <- table(c(as.character(all.network$index1), as.character(all.network$index2)))
TBH.nodes <- match(names(TBH.edges), rownames(TBH))
TBH.conn <- TBH[TBH.nodes, ]
TBH.metric <- colSums(sweep(TBH.conn, 1, TBH.edges, "*"))
TBH.dates <- extract_date(colnames(TBH))

NSH.edges <- table(c(as.character(all.network$index1), as.character(all.network$index2)))
NSH.nodes <- match(names(NSH.edges), rownames(NSH))
NSH.conn <- NSH[NSH.nodes, ]
NSH.metric <- colSums(sweep(NSH.conn, 1, NSH.edges, "*"))
NSH.dates <- extract_date(colnames(NSH))

MAH.edges <- table(c(as.character(all.network$index1), as.character(all.network$index2)))
MAH.nodes <- match(names(MAH.edges), rownames(MAH))
MAH.conn <- MAH[MAH.nodes, ]
MAH.metric <- colSums(sweep(MAH.conn, 1, MAH.edges, "*"))
MAH.dates <- extract_date(colnames(MAH))

CBH.edges <- table(c(as.character(all.network$index1), as.character(all.network$index2)))
CBH.nodes <- match(names(CBH.edges), rownames(CBH))
CBH.conn <- CBH[CBH.nodes, ]
CBH.metric <- colSums(sweep(CBH.conn, 1, CBH.edges, "*"))
CBH.dates <- extract_date(colnames(CBH))

SSH.edges <- table(c(as.character(all.network$index1), as.character(all.network$index2)))
SSH.nodes <- match(names(SSH.edges), rownames(SSH))
SSH.conn <- SSH[SSH.nodes, ]
SSH.metric <- colSums(sweep(SSH.conn, 1, SSH.edges, "*"))
SSH.dates <- extract_date(colnames(SSH))

all.metric <- c(TBH.metric, NSH.metric, MAH.metric, CBH.metric, SSH.metric)
all.dates <- c(TBH.dates, NSH.dates, MAH.dates, CBH.dates, SSH.dates)
lakekey <- c(rep("TBH", length(TBH.metric)), rep("NSH", length(NSH.metric)), rep("MAH", length(MAH.metric)), rep("CBH", length(CBH.metric)), rep("SSH", length(SSH.metric)))
plot.conn <- data.frame(lakekey, all.dates, all.metric)
colnames(plot.conn) <- c("Lake", "Date", "Connectivity")
plot.conn$Lake <- factor(plot.conn$Lake, levels = c("CBH", "NSH", "TBH", "SSH", "MAH"))
ggplot(data = plot.conn, aes(x = Lake, y = Connectivity)) + geom_boxplot() + scale_y_log10() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black"), axis.text = element_text(colour = "black"))

#Test significant differences
lakeids <- levels(plot.conn$Lake)
Lake1 <- character(0)
Lake2 <- character(0)
Test <- character(0)
pvalue <- numeric(0)
Interpretation <- character(0)

for (id in lakeids) { 
  x = subset(plot.conn, plot.conn$Lake == id)
  for (lake in lakeids) {
    y = subset(plot.conn, plot.conn$Lake == lake)
    result <- wilcox.exact(x$Connectivity,y$Connectivity, alternative= "two.sided", conf.level = 0.95)
    Lake1 <- c(Lake1, id)
    Lake2 <- c(Lake2, lake)
    Test <- c(Test, result$method)
    pvalue <- c(pvalue, result$p.value)
    if (result$p.value < 0.05) {
      Interpretation <- c(Interpretation, "Reject Null (difference in median detected)")
    } else {
      Interpretation <- c(Interpretation, "Accept Null (difference in median undetected)")
    }  
  }
}

Wilcoxon.conn <- data.frame(Lake1, Lake2, Test, pvalue, Interpretation)

```

####Figure S6. Connectivity between hypolimnia. Connectivity was measured as the number of edges connecting each OTU multiplied by the abundance of that OTU in a given sample, then summed for the given sample. When the amount of connectivity is compared for lakes with multiple years of sampling, Mary Lake has higer connectivity, followed by South Sparkling Bog. The only lakes that are not significantly different are CBH and TBH.
***

```{r, hclust of all bogs, echo=FALSE}
CBH <- bog_subset("CBH", otu_table)
CBH <- CBH[which(rowSums(CBH) > 75), ]
input <- remove_reps(CBH)
month <- substr(colnames(input), start = 6, stop = 8)
input <- input[, which(month == "MAY" | month == "JUN" | month == "AUG" | month == "JUL" | month == "SEP")]
colnames(input) <- substr(colnames(input), start = 4, stop = 10)
h <- hclust(vegdist(t(input), method = "bray"), method = "ward")

ddata <- dendro_data(h, type = "rectangle")
ddata$year <- factor(substr(ddata$label$label, start = 6, stop = 7), levels = c("05", "07", "08", "09"))

ggplot(segment(ddata)) + geom_segment(aes(x = x, y = y, xend = xend, yend = yend)) + theme(axis.text.x = element_text(colour = "black"), axis.text.y = element_text(colour = "black", angle = 90)) + geom_text(data = label(ddata), aes(x = x, y = y, label = label, colour = ddata$year), vjust = 0.5, hjust = 1.1, size = 5, angle = 90)   + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black"), axis.ticks = element_line(colour="black"), legend.position = "none", axis.text.y = element_text(angle = 0)) + scale_color_manual(values = c("chartreuse4", "darkslategrey", "darkred")) + coord_cartesian(ylim = c(-2, 4)) + labs(title = "Crystal Bog")

NSH <- bog_subset("NSH", otu_table)
NSH <- NSH[which(rowSums(NSH) > 75), ]
input <- remove_reps(NSH)
month <- substr(colnames(input), start = 6, stop = 8)
input <- input[, which(month == "MAY" | month == "JUN" | month == "AUG" | month == "JUL" | month == "SEP")]
colnames(input) <- substr(colnames(input), start = 4, stop = 10)
h <- hclust(vegdist(t(input), method = "bray"), method = "ward")

ddata <- dendro_data(h, type = "rectangle")
ddata$year <- factor(substr(ddata$label$label, start = 6, stop = 7), levels = c("05", "07", "08", "09"))

ggplot(segment(ddata)) + geom_segment(aes(x = x, y = y, xend = xend, yend = yend)) + theme(axis.text.x = element_text(colour = "black"), axis.text.y = element_text(colour = "black", angle = 90)) + geom_text(data = label(ddata), aes(x = x, y = y, label = label, colour = ddata$year), vjust = 0.5, hjust = 1.1, size = 3, angle = 90)   + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black"), axis.ticks = element_line(colour="black"), legend.position = "none", axis.text.y = element_text(angle = 0)) + scale_color_manual(values = c("chartreuse4", "darkslategrey", "darkred")) + coord_cartesian(ylim = c(-2, 6.5)) + labs(title = "North Sparkling Bog")

SSH <- bog_subset("SSH", otu_table)
SSH <- SSH[which(rowSums(SSH) > 75), ]
input <- remove_reps(SSH)
month <- substr(colnames(input), start = 6, stop = 8)
input <- input[, which(month == "MAY" | month == "JUN" | month == "AUG" | month == "JUL" | month == "SEP")]
colnames(input) <- substr(colnames(input), start = 4, stop = 10)
h <- hclust(vegdist(t(input), method = "bray"), method = "ward")

ddata <- dendro_data(h, type = "rectangle")
ddata$year <- factor(substr(ddata$label$label, start = 6, stop = 7), levels = c("05", "07", "08", "09"))

ggplot(segment(ddata)) + geom_segment(aes(x = x, y = y, xend = xend, yend = yend)) + theme(axis.text.x = element_text(colour = "black"), axis.text.y = element_text(colour = "black", angle = 90)) + geom_text(data = label(ddata), aes(x = x, y = y, label = label, colour = ddata$year), vjust = 0.5, hjust = 1.1, size = 3, angle = 90)   + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black"), axis.ticks = element_line(colour="black"), legend.position = "none", axis.text.y = element_text(angle = 0)) + scale_color_manual(values = c("chartreuse4", "darkslategrey", "darkred")) + coord_cartesian(ylim = c(-1.5, 3)) + labs(title = "South Sparkling Bog")

```
####Figure S7. Hierarchical clustering of samples over years. This analysis was performed on hypolimnia with multiple years of sampling. All hypolimnia except for that of Mary Lake show strong clustering patterns by year (see main figures)

***


